\setActionItemPrefix{S-}


\section{Introduction}
This is the Rubin information and data security plan for operations.
This renders \citeds{LDM-324} obsolete.

This security plan conforms with the construction era security documentation:
\begin{itemize}
\item “LSST Information Classification Policy” \citedsp{LPM-122}
\item “LSST Master Information Security Policy” \citedsp{LPM-121}
\end{itemize}

Note has been taken of \cite{TCISSG}.

During the ramp up to operations further security requirements were given to the project by the agencies.
The response to these is in \gls{DMTN}-199\footnote{This is a limited availability document}.

The \gls{RDM} department is concerned with the operations and maintenance  of Rubin constructed software, hardware and networks.
Development remains open-source in nature.

Data from the telescope on Cerro Pach\'{o}n is transmitted over Rubin controlled networks to \gls{SLAC} in  Menlo Park USA.
After an agreed embargo period all data will be transmitted to the \gls{IN2P3} in Lyon, France, and at least some data is transmitted to the Royal Observatory Edinburgh.
Processing occurs in all three locations using a coordinate release of the Rubin \gls{Science Pipelines}.

Access to data will be provided to authorized via a US \gls{DAC}
A data access center similar to the \gls{US} \gls{DAC} will also be provided in Chile.

For data taken initially we have been requested to comply with \citeds{NIST.SP.800-171} the compliance matrix is provided in \appref{ap:compliance}


This document does not do a detailed threat analysis though this should be and are done for individual systems and \gls{Enclave}s. \cite{TCISSG} is good guide for software threats.

\section{Presence of controlled information}\label{sec:cui}
\VRO is an open source project with no confidentiality requirements on the software.
The software project’s integrity requirements are met through the combination of
processes and controls which provide verified user access and protected credentials.
The majority of software testing is conducted using open simulated and observational data sets.

Rubin observatory data is not considered \gls{CUI}, however we have been asked to embargo data for eighty hours and up to ten days for some images.
All data, after the embargo period,  is classified as Internal for two years before it becomes public.
Internal here means available to data rights holders as per \citeds{RDO-013}.
Data rights holders are instructed not to disseminate data outside of the collaboration.
This is the customary protection for this data in the field of optical astronomy.

\input{cyber}
\input{system}
\input{dparc}
\input{tsarc}
\input{rdo}
