\section{System description} \label{sec:desc}

\VRO produces around 20TB of astronomical data per night for the 10 year \gls{LSST}
The control of the the observatory is part of Telescope and site and is covered in \secref{sec:tsarc}.
Data are processed in \gls{SLAC}, \gls{IN2P3} and \gls{ROE}.
A publicly available alert stream emanates from the \gls{USDF} at \gls{SLAC}.
Processing is the responsibility's of \gls{RDP} department while \gls{Quality Assurance} is carried out under \gls{RPF} department.

The Rubin \gls{Operations} Plan \citeds{RDO-018} gives more details.
The \gls{USDF} specification is in \citeds{DMTN-189} but some architecture details are provided in \secref{sec:dparc}.






\subsection{Data verification and quality assurance }
Members of the \gls{RPF} department assess the data quality at short and long timescales.
They require access to tools installed at the summit as well as in the data facilities.
In particular they will need access to images as they are processed to generate alerts.

These team members will have to have \gls{SLAC} accounts to perform this work.
\gls{SLAC} accounts are governed by SLAC policies\footnote{\url{https://policies.slac.stanford.edu}} and
\gls{SLAC} IT specific policies\footnote{\url{https://policies.slac.stanford.edu/categories/information-technology}}.

In addition they will need summit accounts as governed by \citeds{ittn-045} and \citeds{ittn-010}.



\subsection{Data processing}
\subsubsection{Prompt  Processing}
Prompt processing is performed on embargoed data with in SLAC's prompt processing enclave \secref{sec:promptenc}.
This data is subject to \citep{ACP}.

\subsubsection{Data Release Processing}
DRP is performed only on unembargoed data which is fully available to Rubin data rights holders.
Processing is carried out at three sites.
Each site has its own security policies:

\begin{itemize}
\item SLAC lists cyber policies on \url{https://it.slac.stanford.edu/cybersecurity/compliance}.
\item UKDF falls under the \gls{IRIS} security policies listed on \url{https://www.iris.ac.uk/security/}.
\item FrDF lists cyber policies and compliance on \url{https://doc.lsst.eu/cybersecurity/cybersecurity.html}.
\end{itemize}



\subsection{Data access}
After a short embargo period image and derived data are made available to all
data rights holders as defined in \citeds{LDO-13}.

Front end data access via the \gls{RSP} will be hosted on a \gls{cloud} provider such as Google.
Thus not requiring community science users to have \gls{SLAC} accounts.

There will  be a \gls{DAC} in Chile to support Chilean users.
The \gls{UK} intends to host a DAC for \gls{UK} users.

In addition there will be a set of \gls{IDAC}s which will usually serve a portion of the data e.g. perhaps only catalogs or only the object catalog.


\subsection{Rubin \gls{Director}'s Office}
Rubin headquarters are in Tucson Arizona where a few services are also deployed such as Jira and Confluence.
These are discussed more in \secref{sec:rdo}

